Effect of quality-adjusted production function inputs

### 2015-06-18 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Now including CST models.

  modelsCST <- list()
  # Output elasticities based on cost share theorem
  alphaCST <- 0.3 # Fraction of national accounts spent on capital
  betaCST <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    form <- formulas[[country]][["unadjusted"]]
    form <- keep_summands(form, n=-3) # Remove energy
    yVar <- as.name(summands(form, n=1, right=FALSE))
    kVar <- as.name(summands(form, n=1))
    lVar <- as.name(summands(form, n=2))
    tVar <- as.name(summands(form, n=3))
    mod <- lm(log(eval(yVar)) - alpha_CST*log(eval(kVar)) - beta_CST*log(eval(lVar)) ~ eval(tVar), 
              data = historicalData[[country]])
    attr(mod, "naturalCoeffs") <- c(lambda=coef(mod)[[2]],
                                    logscale=coef(mod)[[1]], 
                                    scale=exp(coef(mod)[[1]]),
                                    alpha_1=alphaCST,
                                    alpha_2=betaCST, 
                                    alpha_3=0)
    modelsCST[[country]][["unadjusted"]][["noE"]][["CD"]] <- mod
  }

@@

  #
  # Models where all parameters are estimated
  #

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      nestsWithE <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nestsWithE, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  saveRDS(models, file=fileName)
}
# Put all the coefficients into a data frame
AllCoeffs <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  Country <- id_list[[1]]
  flavor <- id_list[[2]]
  energy <- id_list[[3]]
  model <- id_list[[4]]
  nest <- if(length(id_list) > 4) id_list[[5]] else NA
  res <- rbind(list(Country = Country, 
                    flavor = if (flavor=="unadjusted") "Unadjusted" else "Quality-adjusted",
                    energy = if (energy=="noE") "Without energy" else "With energy",
                    model=model, 
                    nest=nest))
  coeffs <- naturalCoef(mod)
  res <- cbind(res, coeffs)
  # Now calculate the mean squared error (mse) for each model.
  sqResids <- resid(mod)^2
  mse <- sum(sqResids)/(length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
# Get the unadjusted, no-energy values for mse and lambda and place them on each appropriate row
AllCoeffs_unadj_noE <- filter(AllCoeffs, flavor=="Unadjusted", energy=="Without energy") %>% 
                              select(Country, flavor, energy, model, nest, mse, lambda)
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="mse"] <- "mse_unadj_noE"
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="lambda"] <- "lambda_unadj_noE"
AllCoeffs <- merge(AllCoeffs, AllCoeffs_unadj_noE[c("Country", "model", "mse_unadj_noE", "lambda_unadj_noE")])
# Calculate differences relative to the unadjusted, no-energy condition
AllCoeffs <- mutate(AllCoeffs, Dlambda = abs(lambda) - abs(lambda_unadj_noE), 
                    Dmse = mse - mse_unadj_noE,
                    percDlambda = Dlambda / abs(lambda_unadj_noE) * 100)

# Relevel factors
AllCoeffs$Country <- relevelFactor(as.factor(unlist(AllCoeffs$Country)), c("PT", "UK"))
AllCoeffs$flavor <- relevelFactor(as.factor(unlist(AllCoeffs$flavor)), c("Unadjusted", "Quality-adjusted"))
AllCoeffs$energy <- relevelFactor(as.factor(unlist(AllCoeffs$energy)), c("Without energy", "With energy"))
AllCoeffs$model <- relevelFactor(as.factor(unlist(AllCoeffs$model)), c("CD", "CES"))
AllCoeffs$nest <- relevelFactor(as.factor(unlist(AllCoeffs$nest)), c("kl", "kle", "lek", "ekl"))
# Rearrange for easier viewing
AllCoeffs <- arrange(AllCoeffs, Country, model, flavor, energy, nest)
@

>>>>

  modelsCST <- list()
  # Output elasticities based on cost share theorem
  alphaCST <- 0.3 # Fraction of national accounts spent on capital
  betaCST <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    form <- keep_summands(formulas[[country]][["unadjusted"]], n=-3) # Remove energy
    yVar <- as.name(summands(form, n=1, right=FALSE))
    kVar <- as.name(summands(form, n=1))
    lVar <- as.name(summands(form, n=2))
    tVar <- as.name(summands(form, n=3))
    # Cobb-Douglas
    cdModCST <- lm(log(eval(yVar)) - alphaCST*log(eval(kVar)) - betaCST*log(eval(lVar)) ~ eval(tVar), 
                   data = historicalData[[country]])
    attr(cdModCST, "naturalCoeffs") <- data.frame(lambda=coef(cdModCST)[[2]],
                                                  logscale=coef(cdModCST)[[1]], 
                                                  scale=exp(coef(cdModCST)[[1]]),
                                                  alpha_1=alphaCST,
                                                  alpha_2=betaCST, 
                                                  alpha_3=0)
    modelsCST[[country]][["unadjusted"]][["noE"]][["CD"]] <- cdModCST
    # CES
    cesModCST <- plm(log(eval(yVar) / (alphaCST*eval(kVar)^(-rho_1) + betaCST*eval(lVar)^(-rho_1))^(-1/rho_1)) ~ eval(tVar),
                     data = historicalData[[country]],
                     params=c(rho_1=0.8))
    names(cesModCST$coefficients) <- c("logscale", "lambda", "rho_1")
    cesModCST$coefficients <- c(coef(cesModCST), c(delta_1=alphaCST, delta=1))
    attr(cesModCST, "naturalCoeffs") <- EconModels:::makeNatCoef(cesModCST)
    modelsCST[[country]][["unadjusted"]][["noE"]][["CES"]][["kl"]] <- cesModCST
  }

@@

  #
  # Models where all parameters are estimated (models don't adhere to the cost share theorem)
  #

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      nestsWithE <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nestsWithE, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  saveRDS(models, file=fileName)
}
# Put all the cost-share theorem (CST) coefficients into a data frame
AllCoeffsCST <- leaf_apply(modelsCST, class = c("lm"), f = function(modCST, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  res <- rbind(list(Country = id_list[[1]], 
                    flavor = if (id_list[[2]]=="unadjusted") "Unadjusted" else "Quality-adjusted",
                    energy = if (id_list[[3]]=="noE") "Without energy" else "With energy",
                    model=id_list[[4]], 
                    nest=if(length(id_list) > 4) id_list[[5]] else NA,
                    cst=TRUE))
  res <- cbind(res, naturalCoef(modCST))
  # Now calculate the mean squared error (mse) for each model.
  sqResids <- resid(modCST)^2
  mse <- sum(sqResids)/(length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)

AllCoeffsNoCST <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  res <- rbind(list(Country = id_list[[1]], 
                    flavor = if (id_list[[2]]=="unadjusted") "Unadjusted" else "Quality-adjusted",
                    energy = if (id_list[[3]]=="noE") "Without energy" else "With energy",
                    model=id_list[[4]], 
                    nest=if(length(id_list) > 4) id_list[[5]] else NA,
                    cst=FALSE))
  res <- cbind(res, naturalCoef(mod))
  # Now calculate the mean squared error (mse) for each model.
  sqResids <- resid(mod)^2
  mse <- sum(sqResids)/(length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
AllCoeffs <- plyr::rbind.fill(AllCoeffsCST, AllCoeffsNoCST)
# Get reference values for lambda and sse.
ref <- filter(AllCoeffs, cst==TRUE & flavor=="Unadjusted" & energy=="Without energy") %>% 
              select(Country, flavor, energy, model, nest, mse, lambda)

colnames(ref)[colnames(ref)=="mse"] <- "mse_ref"
colnames(ref)[colnames(ref)=="lambda"] <- "lambda_ref"
AllCoeffs <- merge(AllCoeffs, ref[c("Country", "model", "mse_ref", "lambda_ref")])
# Calculate differences relative to the unadjusted, no-energy condition
AllCoeffs <- mutate(AllCoeffs, Dlambda = abs(lambda) - abs(lambda_ref), 
                    Dmse = mse - mse_ref,
                    percDlambda = Dlambda / abs(lambda_ref) * 100)

# Relevel factors
AllCoeffs$Country <- relevelFactor(as.factor(unlist(AllCoeffs$Country)), c("PT", "UK"))
AllCoeffs$flavor <- relevelFactor(as.factor(unlist(AllCoeffs$flavor)), c("Unadjusted", "Quality-adjusted"))
AllCoeffs$energy <- relevelFactor(as.factor(unlist(AllCoeffs$energy)), c("Without energy", "With energy"))
AllCoeffs$model <- relevelFactor(as.factor(unlist(AllCoeffs$model)), c("CD", "CES"))
AllCoeffs$nest <- relevelFactor(as.factor(unlist(AllCoeffs$nest)), c("kl", "kle", "lek", "ekl"))
AllCoeffs$cst <- relevelFactor(as.factor(unlist(AllCoeffs$cst)), c("TRUE", "FALSE"))
# Rearrange for easier viewing
AllCoeffs <- arrange(AllCoeffs, Country, model, flavor, energy, nest, cst)
@
